---
title: "Generating triangular distribution via inverse cdf and writing equations in rmarkdown"
author: "Alex Shen"
date: "2023-03-06"
output:
  html_document:
    theme: cerulean
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Triangular distribution

Triangular distribution is the sum of two independent uniform distribution, with a piece wise pdf of
$$ f(x)= \left\{
\begin{array}{ll}
 x & 0\le x \le 1\\
 2-x & 1 \le x \le 2\\
 0 &otherwise\\
 \end{array}
 \right. $$

and a piece wise CDF of
 
 - $F(x) = \frac{x^2}{2}$, for $0 \le x < 1$;
 - $F(x) = -2+2x-\frac{x^2}{2}$, for $1 \le x \le 2$

Let F(x)=u, and solve x in terms of u to get the inverse CDF
 
 
 - $x = \sqrt{2u}$, for $0 \le u < 1/2$;
 - $x = 2-\sqrt{2(1-u)}$, for $1/2 \le u \le 1$
 
 
A few other Latex like formulas, please refer to

- [R markdown for Scientist](https://rmd4sci.njtierney.com/math)
- [Arthoring books with R Markdown](https://bookdown.org/yihui/bookdown/markdown-extensions-by-bookdown.html)
- [Mathematics in R markdown](https://rpruim.github.io/s341/S19/from-class/MathinRmd.html)
- [An example R Markdown](https://www.math.mcgill.ca/yyang/regression/RMarkdown/example.html)
$$
\sum_{i = 1}^{n}{(\bar{x} - x_i)^2}
$$

$$
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
$$

$$
Pr(\theta | y) \propto Pr(y | \theta) Pr(\theta)
$$

$$
Y \sim X\beta_0 + X\beta_1 + \epsilon
$$

$$
\epsilon \sim N(0,\sigma^2)
$$

$$f(k) = {n \choose k} p^{k} (1-p)^{n-k}$$

$$X = \begin{bmatrix}1 & x_{1}\\
1 & x_{2}\\
1 & x_{3}
\end{bmatrix}$$

\begin{equation*} 
\frac{d}{dx}\left( \int_{a}^{x} f(u)\,du\right)=f(x)
\end{equation*}

\begin{equation} 
\begin{split}
\mathrm{Var}(\hat{\beta}) & =\mathrm{Var}((X'X)^{-1}X'y)\\
 & =(X'X)^{-1}X'\mathrm{Var}(y)((X'X)^{-1}X')'\\
 & =(X'X)^{-1}X'\mathrm{Var}(y)X(X'X)^{-1}\\
 & =(X'X)^{-1}X'\sigma^{2}IX(X'X)^{-1}\\
 & =(X'X)^{-1}\sigma^{2}
\end{split}
(\#eq:var-beta)
\end{equation}


\begin{align*}
x & = \frac{-b \pm \sqrt{b^{2} - 4ac}}{2a}\\
a & = b \\
X &\sim {\sf Norm}(10, 3) \\
5 & \le 10
\end{align*}

$$
\widehat{\theta}_1=\frac{X_1+\ldots+X_{n_1}}{n_1}
$$

x=\frac{-b \pm \sqrt{b^2-4 a c}}{2 a}

$x=\frac{-b \pm \sqrt{b^2-4 a c}}{2 a}$

$$
x=\frac{-b \pm \sqrt{b^2-4 a c}}{2 a}
$$

\begin{equation}
x=\frac{-b \pm \sqrt{b^2-4 a c}}{2 a}
\end{equation}

## Generating triangular distribution
```{r}
u <- runif(10^6) #generate U
x <- sqrt(2*u) # create x based on first component
indicator <- u>0.5 #
x[indicator] <- 2-sqrt(2*(1-u[indicator])) #replace x based on 2nd component

hist(x, freq = F)
curve(x^1, from=0, to =1, col = "blue",add =TRUE)
curve(2-x, from =1, to =2, col ="blue", add =T)

```



